---
title: "IRT for beginners"
subtitle: "Visualizzare aiuta"
institute: "Bressanone"
author: "Dr. Ottavia M. Epifania"
date: "Corso IRT @ Università Libera di Bolzano, \\ 17-18 Gennaio 2023"
output: ioslides_presentation
runtime: shiny
---
<style type="text/css">


code.r{
  font-size: 14px;
}
pre {
  font-size: 12.5px
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff=80), tidy= TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA, 
  echo = FALSE, 
  warning = FALSE
)
library(ggplot2)
library(sjPlot)
library(corrplot)
library(nnet)
library(formatR)
library(kableExtra)
library(shiny)

IRT <- function(theta, a = 1, b = 0, c = 0,e = 1) {
  y <- c + (e - c) * exp(a * (theta - b)) / (1 + exp(a * (theta - b)))
  y[is.na(y)] = 1
  return(y)
}

i_info <- function(b, a=1,c=0, theta = seq(-5,5,length.out=1000)){
 
P <- NULL 
Q <- NULL
Ii <- NULL
for(i in 1:1000){
  P[i] <- 1/(1+ exp (-a*(theta[i] - b)))
  Q[i]= 1-P[i]
  Ii[i] =(a*Q[i]*(P[i]-c)^2)/(P[i]*((1-c)^2)) # (3PL)
   }
return(Ii)
}
# Function to get all item information
item_info <- function(b,a=1, c= 0){
item <- NULL
  for(i in 1:length(b)){
  item[[i]] <- i_info(b[i],a[i])
  }
return(item)
}

set.seed(999)
```



## Tutti quanti

```{r}
shiny::shinyApp(
  ui <- fluidPage(



  # Sidebar layout with input and output definitions ----
  verticalLayout(
      # App title ----

    # Sidebar to demonstrate various slider options ----
    wellPanel(


      # Input: Decimal interval with step value ----
      fluidRow(
        column(6, div(style="height: 20px;",
                      sliderInput("b", "Difficulty:",
                  min = -5, max = 5,
                  value = 0, step = 0.1))),
              column(6,   sliderInput("lucky", "Lucky guess:",
                  min = 0, max = 0.5,
                  value = 0, step = 0.01))),
      fluidRow(
        column(6,    sliderInput("a", "Discrimination:",
                  min = -5, max = 5,
                  value = 1, step = 0.1)),
                column(6,  sliderInput("care", "Careless error:",
                  min = 0.5, max = 1,
                  value = 1, step = 0.01)))



    ),


      plotOutput("plot", height = "250px")
  )
),


server <- function(input, output) {
IRT <- function(theta, a = 1, b = 0, c = 0,e = 1) {
  y <- c + (e - c) * exp(a * (theta - b)) / (1 + exp(a * (theta - b)))
  y[is.na(y)] = 1
  return(y)
}


 output$plot <- renderPlot({
   theta <- theta <- seq(-7, 7, .001)
  par(mar = c(5,7,4,2) + 0.1)

 plot(theta, IRT(theta, b = 0, a = 1),
     cex.lab= 2,
     cex.axis =1.5,
       xlab = expression(theta), ylab = bquote(.("P(x = 1)")),
       xlim = c(-5, 5), ylim = c(0, 1),
     type = "l", lwd = 1, lty = 2,
     col = "royalblue")
 lines(theta, IRT(theta, b=input$b,
                 a=input$a,
                 c= input$lucky,
                 e= input$care),
       type = "l", lwd = 3,
     col = "seagreen")
  })

},
options = list(height = 400)

)

```


## In pratica: ICC, IIF e TIF

```{r}
shiny::shinyApp(
  ui <- fluidPage(
tags$style(HTML(".js-irs-0 .irs-single, .js-irs-0 .irs-bar-edge, .js-irs-0 .irs-bar {background: royalblue}")),
  tags$style(HTML(".js-irs-1 .irs-single, .js-irs-1 .irs-bar-edge, .js-irs-1 .irs-bar {background: magenta}")),
  tags$style(HTML(".js-irs-2 .irs-single, .js-irs-2 .irs-bar-edge, .js-irs-2 .irs-bar {background: seagreen}")),
tags$style(HTML(".js-irs-3 .irs-single, .js-irs-3 .irs-bar-edge, .js-irs-3 .irs-bar {background: royalblue}")),
  tags$style(HTML(".js-irs-4 .irs-single, .js-irs-3 .irs-bar-edge, .js-irs-4 .irs-bar {background: magenta}")),
  tags$style(HTML(".js-irs-5 .irs-single, .js-irs-5 .irs-bar-edge, .js-irs-5 .irs-bar {background: seagreen}")),
  # Sidebar layout with input and output definitions ----
  verticalLayout(
      # App title ----
    # Sidebar to demonstrate various slider options ----
    wellPanel(
      # Input: Decimal interval with step value ----
      fluidRow(
        column(4, div(style="height: 20px;", 
                      sliderInput("b1", "b1",
                  min = -3, max = 3,
                  value = 0, step = 0.1))),
              column(4,   sliderInput("b2", "b2",
                  min = -3, max = 3,
                  value = 0, step = 0.1)),
        column(4,   sliderInput("b3", "b3",
                  min = -3, max = 3,
                  value = 0, step = 0.1))), 
      fluidRow( 
        column(4,    sliderInput("a1", "a1",
                  min = 0.20, max = 3,
                  value = 1, step = 0.1)), 
                column(4,  sliderInput("a2", "a2",
                  min = 0, max = 3,
                  value = 0.70, step = 0.1)), 
        column(4,    sliderInput("a3", "a3",
                  min = 0, max = 3,
                  value = 1.90, step = 0.1)))
    ),
fluidRow(
  column(6,
         div(style = "height=20px", 
             plotOutput("plot", height = "250px"))), 
  column(6, 
         div(style = "height=20px", 
             plotOutput("tif", height = "250px")))
)
  )
),
server <- function(input, output) {
IRT <- function(theta, a = 1, b = 0, c = 0,e = 1) {
  y <- c + (e - c) * exp(a * (theta - b)) / (1 + exp(a * (theta - b)))
  y[is.na(y)] = 1
  return(y)
}
i_info <- function(b, a=1,c=0, theta = seq(-5,5,length.out=1000)){
 
P <- NULL 
Q <- NULL
Ii <- NULL
for(i in 1:1000){
  P[i] <- 1/(1+ exp (-a*(theta[i] - b)))
  Q[i]= 1-P[i]
  Ii[i] =(a*Q[i]*(P[i]-c)^2)/(P[i]*((1-c)^2)) # (3PL)
   }
return(Ii)
}
# Function to get all item information
item_info <- function(b,a=1){
item <- NULL
  for(i in 1:length(b)){
  item[[i]] <- i_info(b[i],a[i])
  }
return(item)
}
 output$plot <- renderPlot({
   theta <- theta <- seq(-7, 7, .001)
  par(mar = c(5,7,4,2) + 0.1)
  b = c(input$b1, input$b2, input$b3)
  a = c(input$a1, input$a2, input$a3)
c <- item_info(b,a)
Theta <- matrix(seq(-4,4, length.out=1000))
check <- data.frame(Theta,
                    item_info = c[[1]],
                    item_info2 = c[[2]],
                    item_info3 = c[[3]])
d1 <- do.call('cbind',c)
sum_info2 <- rowSums(d1)
plot(check$Theta, check$item_info, cex.lab= 2, main = "IIF",
      cex.axis =1.5,
        xlab = expression(theta), ylab = expression(paste("I(", theta, ")")),
 type = "l", lwd =2,
      col = "royalblue", ylim = c(0,1))
 lines(check$Theta, check$item_info2, lwd =2,
       col = "magenta", lty = 4)
 lines(check$Theta, check$item_info3, lwd =2,
       col = "seagreen", lty = 2)
 
# icc degli stessi item 
lines(theta, IRT(theta, b=input$b1,
                 a=input$a1),
       type = "l", lwd = 1, lty = 2,
     col = "royalblue")
 lines(theta, IRT(theta, b=input$b2,
                 a=input$a2),
       type = "l", lwd = 1, lty = 2,
     col = "magenta")
 lines(theta, IRT(theta, b=input$b3,
                 a=input$a3),
       type = "l", lwd = 1, lty = 2,
     col = "seagreen")
  })
  
 
 output$tif <- renderPlot({
   theta <- theta <- seq(-7, 7, .001)
  par(mar = c(5,7,4,2) + 0.1)
  b = c(input$b1, input$b2, input$b3)
  a = c(input$a1, input$a2, input$a3)
c <- item_info(b,a)
Theta <- matrix(seq(-4,4, length.out=1000))
check <- data.frame(Theta,
                    item_info = c[[1]],
                    item_info2 = c[[2]],
                    item_info3 = c[[3]])
d <- do.call('cbind',c)
sum_info1 <- rowSums(d)
d1 <- do.call('cbind',c)
sum_info2 <- rowSums(d1)
   
Theta <- matrix(seq(-4,4, length.out=1000))
check <- data.frame(Theta, sum_info1, sum_info2)
par(mar = c(5, 4, 4, 4) + 0.3)  # Leave space for z axis
plot(check$Theta, check$sum_info2,
       type = "l", lwd =2,
     col = "black", ylim = c(0, 2),
       xlab = expression(theta), 
     ylab = expression(paste("I(", theta, ")")), cex.lab= 2, 
     cex.axis=1.5, main = "TIF")
 })
 # par(new = TRUE)
 # plot(check$Theta, sqrt(1/check$sum_info2), cex.lab= 2, 
 #     cex.axis =1.5,
 #       xlab = expression(theta), ylab = "SEM", 
 #     type = "l", lwd =2, 
 #     col = "firebrick")
},      
options = list(height = 400)    
)
```





# Oltre il dicotomico 

## Polytomous Rasch Model 

$P(X = x, x > 0| \theta_s, \tau_{ic}) = \frac{exp \left(\sum_{c = 1}^{m} \theta_s - \tau_{ic}\right)}{1 + exp \left(\sum_{c = 1}^{m} \theta_s - \tau_{ic} \right)}$

```{r, warning = FALSE, message=FALSE,  eval = TRUE, fig.align='center'}
library(psych)
library(ltm)
set.seed(1234)
data <- sim.poly(nvar = 4, n = 500, 
                     low = -1, high = 2, mod = "logistic", 
                     cat = 4, a = c(0.2, 0.7, 1.20, 1.90))
prm <- gpcm(data$items, constraint = "rasch")
par(mfrow=c(2,2))
plot(prm, type = "ICC")
```

## Partial credit model

$P(X = x, x > 0| \theta_s, \tau_{ic}, a) = \frac{exp \left[\sum_{c = 1}^{m} a (\theta_s - \tau_{ic})\right]}{1 + exp \left[\sum_{c = 1}^{m} a (\theta_s - \tau_{ic})\right]}$

```{r fig.align='center'}
pcm <- gpcm(data$items, constraint = "1PL")
par(mfrow=c(2,2))
plot(pcm, type = "ICC")
```

## Graded Response Model 

$P(X = x, x > 0| \theta_s, \tau_{ic}, a_i) = \frac{exp \left[\sum_{c = 1}^{m} a_i (\theta_s - \tau_{ic})\right]}{1 + exp \left[\sum_{c = 1}^{m} a_i (\theta_s - \tau_{ic})\right]}$

```{r, fig.align='center'}
grm <- gpcm(data$items, constraint = "gpcm")
par(mfrow=c(2,2))
plot(grm, type = "ICC")
```


